---
title: "Microbial Physiology I1"
author: "Leonardo Claudin and Catherine Chia"
date: "18/2/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning=FALSE
  
)
library("dplyr")
library("plyr")
library("tidyverse")
library("ggplot2")
library("ggpubr")
```

Load the settings and the script with the loading functions:

```{r load_settings}
# Load project settings
source(here::here("settings.R"))
# Load helper functions
source(file.path(functions.dir, "output.R"))
```

Load all datasets into different data.frames with the following code:

```{r load_data}
df.turbidostat <- read_output_csv("turbidostat", .output.dir = data.dir)

df.design <- read_output_csv("design", .output.dir = data.dir)

df.od <- read_output_csv("spectrophotometer", .output.dir = data.dir)
```

```{r}
od.design = left_join(df.design, df.od, by = "channel")
```

## Estimating pigment composition.
```{r}
chl.a.comp = data.frame(channel = od.design$channel,
                      conc = 10.186*(od.design$od_value_685-od.design$od_value_730)-0.08,
                      purpose = od.design$purpose)
chl.a.m = data.frame(mean = tapply(chl.a.comp$conc, chl.a.comp$purpose, mean),
                     median = tapply(chl.a.comp$conc, chl.a.comp$purpose, median),
                     IQR = tapply(chl.a.comp$conc, chl.a.comp$purpose, IQR),
                     sd = tapply(chl.a.comp$conc, chl.a.comp$purpose, sd))

PBS.comp = data.frame(channel = od.design$channel,
                      conc = (od.design$od_value_634-od.design$od_value_730)/od.design$od_value_730,
                      purpose = od.design$purpose)

PBS.m = data.frame(mean = tapply(PBS.comp$conc, PBS.comp$purpose, mean),
                   median = tapply(PBS.comp$conc, PBS.comp$purpose, median),
                   IQR = tapply(PBS.comp$conc, PBS.comp$purpose, IQR),
                     sd = tapply(PBS.comp$conc, PBS.comp$purpose, sd))

```


## Statistical test for Chl a and PBS

```{r}
# Split lists according to conditions
ls_chla_treatment <- ifelse(chl.a.comp$purpose=="treatment",chl.a.comp$conc,999)
ls_chla_control <- ifelse(chl.a.comp$purpose=="control",chl.a.comp$conc,999)
ls_pbs_treatment <- ifelse(PBS.comp$purpose=="treatment",PBS.comp$conc,999)
ls_pbs_control <- ifelse(PBS.comp$purpose=="control",PBS.comp$conc,999)

# Remove the nonconformities
ls_chla_treatment[ ls_chla_treatment != 999]
ls_chla_control[ ls_chla_control != 999]
ls_pbs_treatment[ ls_pbs_treatment != 999]
ls_pbs_control[ ls_pbs_control != 999]

#Normality test. p<0.05
shapiro.test(ls_chla_treatment) #p-value = 0.004042, non-normally distributed
shapiro.test(ls_chla_control) #p-value = 0.004043, non-normally distributed
shapiro.test(ls_pbs_treatment) #p-value = 0.004043, non-normally distributed
shapiro.test(ls_pbs_control) #p-value = 0.004043, non-normally distributed

#Wilcoxon
# the p-value is less than α =.05, there is sufficient evidence to reject the null hypothesis, meaning treatment and control are different
# one-tailed test, because we wanna test if less phosphate leads to less ploidy
wilcox.test(ls_chla_treatment, ls_chla_control, alternative = "less") #p-value = 0.2471
wilcox.test(ls_pbs_treatment, ls_pbs_control, alternative = "less") #p-value = 0.3662

```

## Plot pigments
```{r}
pig.plot = rbind(chl.a.m, PBS.m)
rownames(pig.plot) = NULL
pig.plot$pig = c(rep("chl.A", 2), rep("PBS",2))
pig.plot$purpose = c("Control", "Treatment", "Control", "Treatment")

ggplot(data=pig.plot, aes(x=as.factor(pig), y=mean, fill=purpose)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual("legend", values = c("Control" = "#00BBDB", "Treatment" = "#F066EA")) +
  labs(y = "Chl.A + PBS (mg pigment/L)", x = "Chlorophyll A and PBS") +
  guides(fill=guide_legend(title="Group")) +
  ggtitle("Concentration of Chlorophyll a and Phycobilisome in both conditions")+ 
  geom_errorbar(aes(ymin=median-IQR, ymax=median+IQR), width=.2,
                 position=position_dodge(.9)) +
  theme_classic()
```


## Turbidostat Data. 
```{r}
df.turb_f = left_join(df.turbidostat, df.design, by = "channel")

ggplot(df.turb_f, aes(x = time_h, y = OD, color = purpose)) +
  geom_point() +
  facet_wrap(~ channel + purpose) +
  theme_bw() +
  labs(y = expression("OD"[720]),
       x = "time (h)") +
  scale_color_manual("legend", values = c("control" = "#00BBDB", "treatment" = "#F066EA")) +
  theme(legend.position = "none") 
```

## Turbidostat results if we zoom in.
```{r}
df.turb.zoom = df.turb_f %>% filter(df.turb_f$time_h > 280,
                                    df.turb_f$time_h < 320)
ggplot(df.turb.zoom, aes(x = time_h, y = OD, color = pump)) +
  geom_point() +
  facet_wrap(~ channel + purpose) +
  theme_bw() +
  labs(y = expression("OD"[720]),
       x = "time (h)") +
  scale_color_manual("legend", values = c("FALSE" = "#FF689F", "TRUE" = "#00BBDB")) +
  guides(color=guide_legend(title="Pump")) +
  theme(legend.position="bottom")
```

## Determining the growth rates of each cycle.
```{r}
models <- dlply(df.turbidostat, c("channel"), function(df){
  dlply(df, "decision", function(df2){
    model = lm(log(OD) ~ time_h, data = df2)
    coefs = model$coef[2]
    R2 = summary(model)$r.squared
    time = mean(df2$time_h)
    c(coefs, R2, time)
  })
  })

df.turb.zoom2 = df.turb_f %>% filter(df.turb_f$channel %in% c(3,8),
                                    df.turb_f$time_h > 151,
                                    df.turb_f$time_h < 180)

ggplot(df.turb.zoom2, aes(x = time_h, y = log(OD), color = pump)) +
  geom_point() +
  geom_smooth(method = "lm", inherit.aes = F, aes(x = time_h, y = log(OD), fill = as.factor(decision)),
              show.legend = F, formula = "y~x") +
  facet_wrap(~ channel + purpose) +
  theme_bw() +
  labs(y = expression(paste("log(OD"[720],")")),
       x = "time (h)") +
  scale_color_manual("legend", values = c("FALSE" = "#FF689F", "TRUE" = "#00BBDB")) +
  guides(color=guide_legend(title="Pump")) +
  theme(legend.position="bottom")
```


```{r}
lengths = as.vector(do.call(rbind, lapply(models, function(x) length(x))))
channel = unlist(mapply(rep, c(1,3,4,5,7,8), lengths))

df.turb.lm = as.data.frame(cbind(do.call(rbind, lapply(models, function(x){
  cbind(do.call(rbind, x), names(x))}
  )), channel))
rownames(df.turb.lm) = NULL
colnames(df.turb.lm) = c("growth.r", "R2", "time", "decision", "channel")
```

```{r}
df.turb.lm = df.turb.lm %>%
  mutate(purpose = case_when(
    as.character(channel) %in% c(1,3,4) ~ "treatment",
    as.character(channel) %in% c(5,7,8) ~ "control"
    ))
df.turb.lm$growth.r = as.numeric(df.turb.lm$growth.r)
df.turb.lm$R2 = as.numeric(df.turb.lm$R2)
df.turb.lm$time = as.numeric(df.turb.lm$time)
```

```{r}
ggplot(df.turb.lm, aes(x = time, y = growth.r, color = R2)) +
  geom_point() +
  facet_wrap(~ channel + purpose) +
  labs(y = expression(paste("growth rate (h"^-1, ")")),
       x = "time (h)") +
  guides(color=guide_colorbar(title=expression("R"^2))) +
  theme(legend.position = "right") +
  scale_color_gradient(low="yellow", high="blue") +
  theme_bw()
```

```{r}
ggplot(df.turb.lm %>% filter(df.turb.lm$R2>=0.9,
                             df.turb.lm$growth.r<0.2), 
       aes(x = time, y = growth.r, color = R2)) +
  geom_point() +
  facet_wrap(~ channel + purpose) +
  labs(y = expression(paste("growth rate (h"^-1, ")")),
       x = "time (h)") +
  guides(color=guide_colorbar(title=expression("R"^2))) +
  theme(legend.position = "right") +
  scale_color_gradient(low="yellow", high="blue") + 
  theme_bw()
```

### Sanity Check: Correlation between Dilution and Growth Rate.

```{r}
df.turb_f.dil.pump = df.turb_f %>% filter(pump==T)
df.turb_f.dil.first = df.turb_f %>% group_by(channel, decision) %>% slice(1)
df.turb_f.dil = rbind(df.turb_f.dil.first, df.turb_f.dil.pump)
df.turb_f.dil = df.turb_f.dil %>% filter(time_h > 0.1) %>% arrange(channel, decision)
```

```{r}
ggplot(df.turb_f.dil, aes(x = time_h, y = OD, color = pump)) +
  geom_point() +
  facet_wrap(~ channel + purpose) +
  theme_bw() +
  labs(y = expression("OD"[720]),
       x = "time (h)") +
  scale_color_manual("legend", values = c("FALSE" = "#FF689F", "TRUE" = "#00BBDB")) +
  guides(color=guide_legend(title="Pump")) +
  theme(legend.position="bottom")
```

```{r}
df.turb_f.dil.f = df.turb_f.dil %>% group_by(channel) %>% mutate(dil_pair = ceiling(row_number() / 2)) %>% ungroup()
df.turb_f.dil.f = df.turb_f.dil.f %>% group_by(channel, dil_pair) %>%
  mutate(Diff = lag(OD)/OD)
df.turb_f.dil.f = df.turb_f.dil.f %>% filter(! is.na(Diff))
df.turb.lm.dil = df.turb.lm %>% group_by(channel) %>% slice(2:n())
df.turb.dil.plot = cbind(df.turb.lm.dil,Dilution = df.turb_f.dil.f$Diff)
```

```{r}
df.turb.dil.plot.f = df.turb.dil.plot %>% filter(R2>=0.9,growth.r<0.2, Dilution<1.25)
ggplot(df.turb.dil.plot.f,
       aes(x = Dilution, y = growth.r, color = channel)) +
  geom_point(width = 0.7) +
  labs(x = "Dilution Factor", y = expression(paste("growth rate [",mu, "(h"^-1,")]")),
       title = "Sanity Check: Growth rate - Dilution Correlation") +
  theme_bw()
```

```{r}
models.dil <- dlply(df.turb.dil.plot.f, c("channel"), function(df){
  model = lm(growth.r ~ Dilution, data = df)
  summary(model)$r.squared
  })
rsquared.dil = as.data.frame(as.matrix(flatten(models.dil)))
colnames(rsquared.dil) = c("R2")
rsquared.dil$channel = as.double(rownames(rsquared.dil))
rsquared.dil = left_join(rsquared.dil, od.design, by = "channel")
```

```{r}
ggplot(rsquared.dil, aes(x = as.factor(channel), y = as.numeric(R2), fill = purpose)) +
  geom_bar(width = 0.7, stat='identity', position = position_dodge()) + 
  scale_fill_manual("legend", values = c("control" = "#00BBDB", "treatment" = "#F066EA")) +
  theme(legend.position = "none") +
  labs(x = "Channel", y = expression("R"^2), 
       title = expression(paste("R"^2, "value of linear models from each turbidostat channel"))) +
  guides(fill=guide_legend(title="Purpose")) +
    theme_bw()
```


## Summarizing Growth Rates: last 5 days of cultivation.

```{r}
## TODO Verfiy time > 280 with METHODS |Total 400h, get last 120h
df.turb.lm.5 = df.turb.lm %>% filter(df.turb.lm$R2>=0.9,
                             df.turb.lm$growth.r<0.2,
                             df.turb.lm$time >280)
ggplot(df.turb.lm.5,aes(x = time, y = growth.r, color = R2)) +
  geom_point() +
  facet_wrap(~ channel + purpose) +
  labs(y = expression(paste("growth rate (h"^-1, ")")),
       x = "time (h)") +
  guides(color=guide_colorbar(title=expression("R"^2))) +
  theme(legend.position = "right") +
  scale_color_gradient(low="yellow", high="blue") + 
  theme_bw()
```

```{r}
df.turb.msd.5 = data.frame(df.turb.lm.5 %>% 
                             group_by(channel) %>% 
                             summarise_at(vars(growth.r),
                                          list(median)),
                           df.turb.lm.5 %>% 
                             group_by(channel) %>% 
                             summarise_at(vars(growth.r),
                                          list(IQR)))
df.turb.msd.5 = df.turb.msd.5[,-3]
df.turb.msd.5$channel = as.numeric(df.turb.msd.5$channel)
colnames(df.turb.msd.5) = c("channel", "median", "IQR")
df.turb.msd.5 = left_join(df.design, df.turb.msd.5, by = "channel")
```

```{r}
ggplot(data=df.turb.msd.5, aes(x=as.factor(channel), y=median, fill=purpose)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual("legend", values = c("control" = "#00BBDB", "treatment" = "#F066EA")) +
  labs(y = expression(paste(mu, "(h"^-1,")")), x = "Channel",
       title = "Summary of technical variation",
       subtitle = "Bars show median value per channel and error bars indicate IQR") +
  guides(fill=guide_legend(title="Purpose")) +
  geom_errorbar(aes(ymin=median-IQR, ymax=median+IQR), width=.2,
                 position=position_dodge(.9)) +
  theme_classic()
```

```{r}
df.turb.msd.Ph = data.frame(df.turb.msd.5 %>% 
                             group_by(phosphate_mM) %>% 
                             summarise_at(vars(median),
                                          list(median)),
                           df.turb.msd.5 %>% 
                             group_by(phosphate_mM) %>% 
                             summarise_at(vars(median),
                                          list(IQR)))
df.turb.msd.Ph = df.turb.msd.Ph[,-3]
colnames(df.turb.msd.Ph) = c("Phosphate_mM", "median", "IQR")
df.turb.msd.Ph$Phosphate_mM = as.character(df.turb.msd.Ph$Phosphate_mM)
```

```{r}
ggplot(data=df.turb.msd.Ph, aes(x=as.factor(Phosphate_mM), y=median, fill = Phosphate_mM)) +
  geom_bar(stat="identity", position=position_dodge()) +
  labs(y = expression(paste(mu, "(h"^-1,")")), x = "[Phosphate] (mM)",
       title = "Median growth rate per condition") +
  scale_fill_manual("legend", values = c("0.23" = "#00BBDB", "0.0115" = "#F066EA")) +
  geom_errorbar(aes(ymin=median-IQR, ymax=median+IQR), width=.2,
                 position=position_dodge(.9)) +
  theme_classic() +
  theme(legend.position="null")
```

```{r}
ggplot(df.turb.lm.5, aes(x = channel, y = growth.r, fill = purpose)) +
  geom_violin(trim = F) +
  scale_fill_manual("legend", values = c("control" = "#00BBDB", "treatment" = "#F066EA")) +
  labs(y = expression(paste("growth rate (h"^-1, ")")),
       x = "Channel",
       title = "Violin plot",
       subtitle = "Growth rates distribution per channel") +
  geom_point(alpha = 0.3, color = "darkblue") +
  geom_boxplot(width=0.1, alpha = 0.6, fill = "white") +
  theme_classic() +
  theme(legend.position="null",
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1, color="gray" ) ) +
  scale_y_continuous(breaks = round(seq(min(df.turb.lm.5$growth.r), max(df.turb.lm.5$growth.r), by = 0.005),3))
```



### Sanity Check: Normality per condition.

```{r}
wilcox.test(df.turb.msd.5$median ~ df.turb.msd.5$purpose)
```