---
title: "Microbial Physiology II"
author: "Leonardo Claudin and Catherine Chia"
date: "02/03/2022"
output: pdf_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning=FALSE
  
)
library("dplyr")
library("plyr")
library("tidyverse")
library("ggplot2")
library("ggpubr")
library("gridExtra")
```

Load the settings and the script with the loading functions:

```{r load_settings}
# Load project settings
source(here::here("settings.R"))
# Load helper functions
source(file.path(functions.dir, "output.R"))
```

Load all datasets into different data.frames with the following code:

```{r load_data}
df.qpcr <- read_output_csv("qpcr", .output.dir = data.dir)
df.casy <- read_output_csv("casy", .output.dir = data.dir)
#df.turbidostat <- read_output_csv("turbidostat", .output.dir = data.dir)
df.design <- read_output_csv("design", .output.dir = data.dir)
#df.od <- read_output_csv("spectrophotometer", .output.dir = data.dir)
```

#qPCR Results.
## Standard Curve.

```{r}
df.qpcr.st = df.qpcr %>% filter(df.qpcr$sample_class == "standard")
df.qpcr.st2 = df.qpcr.st
df.qpcr.st2$replicate = "all"
df.qpcr.st2 = rbind(df.qpcr.st2, df.qpcr.st)
```


```{r}
MW_DNA = 335077.77
NAv = 6.02214076*(10^23)

ggplot(df.qpcr.st2, aes(x = log10((0.04*template_vol*NAv)/(dilution_factor*final_vol*10^9*MW_DNA)), 
                        y = Ct)) +
  geom_point() +
  geom_smooth(method = "lm", formula = "y~x") +
  facet_wrap(~ replicate) +
  theme_bw() +
  labs(title = "Standard curves.",
       subtitle = expression(paste("Linear regression from DNA fragments of known concentration 0.04ng/",mu,"L",)),
       x = expression(paste("log"[10],"DNA (fragments x volume"^-1, ")")),
       y = "Ct")
```

### Obtaining the Amplification efficiency of standard curves.

```{r}
models = dlply(df.qpcr.st2, c("replicate"), function(df){
  model = lm(Ct ~ log10((0.04*template_vol*NAv)/(dilution_factor*final_vol*10^9*MW_DNA)), data = df)
  coefs = model$coef[2]
  R2 = summary(model)$r.squared
  intercept = model$coefficients[1]
  c(coefs, R2, intercept)
})

df.qpcr.st.c = as.data.frame(do.call(rbind, models))
colnames(df.qpcr.st.c) = c("slope", "R2", "intercept")

df.qpcr.st.c$replicate = rownames(df.qpcr.st.c)
rownames(df.qpcr.st.c) = NULL

df.qpcr.st.c$efficiency = (10^(-1/df.qpcr.st.c$slope))/2*100
```

### Selection of best standard curve.

```{r}
ggplot(df.qpcr.st.c, aes(x = replicate, y = slope)) +
  geom_point() +
  labs(x = "Sample", y = "Slope", title = "Red lines indicate optimal slope") +
  geom_hline(yintercept = -1/log10(2), color = "red") +
  theme_bw()
```

```{r}
ggplot(df.qpcr.st.c, aes(x = replicate, y = efficiency)) +
  geom_point() +
  labs(x = "Sample", y = "Efficiency", title = "Red lines indicate optimal range") +
  geom_hline(yintercept = 90, color = "red") +
  geom_hline(yintercept = 105, color = "red") +
  theme_bw()
```

```{r}
ggplot(df.qpcr.st.c, aes(x = replicate, y = R2)) +
  geom_point() +
  labs(x = "Sample", y = expression("R"^2), title = "Red line indicate minimum R2") +
  geom_hline(yintercept = 0.98, color = "red") +
  theme_bw()
```

## Ploidy Determination.

```{r}
selected.st = df.qpcr.st.c %>% filter(df.qpcr.st.c$replicate==1)
b = selected.st$intercept
a = selected.st$slope

df.qpcr.smpl = df.qpcr %>% filter(df.qpcr$sample_class == "sample")
df.qpcr.smpl$nDNA_fragments = 10^((df.qpcr.smpl$Ct-b)/a)*(df.qpcr.smpl$final_vol/df.qpcr.smpl$template_vol)*df.qpcr.smpl$dilution_factor
```

```{r}
ggplot(df.qpcr.smpl, aes(x = as.factor(channel), 
                         y = nDNA_fragments, 
                         color = as.factor(dilution_factor))) +
  geom_boxplot(color = "black") +
  geom_point(alpha = 0.7) +
  labs(y = expression(paste("[DNA] (# fragments x ",mu,"L"^-1,")")),
       x = "Channel",
       title = "Concentration of DNA fragments per channel") +
  guides(color=guide_legend(title = "Dilution Factor")) +
  theme_bw()
```

# CASY COUNTER

```{r}
df.casy.mc = df.casy %>% filter(df.casy$sample_type=="MC")
df.casy.mc$counts = df.casy.mc$counts*10010/(200*10)*1000
df.casy.mc.m = data.frame(df.casy.mc %>% 
                            group_by(channel) %>%
                            summarise_at(vars(counts),
                                         list(mean, sd)))
colnames(df.casy.mc.m) = c("channel", "mean", "sd")
df.casy.mc.m = left_join(df.design, df.casy.mc.m, by = "channel")
```

```{r}
ggplot(df.casy.mc.m, aes(x = as.factor(channel), y = mean/(10^6), fill = purpose)) +
  geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=(mean-sd)/(10^6), ymax=(mean+sd)/(10^6)), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_manual("legend", values = c("control" = "#00BBDB", "treatment" = "#F066EA")) +
  labs(x = "Channel", y = "Million cells / mL",
       title = "Cell counts in MC samples") +
    guides(fill=guide_legend(title="Purpose")) +
  theme_bw()
```
###Calculate broken cells per channel
# Find the mean per channel per BBB or ABB (removing technical error)

```{r}
df.casy.abb = df.casy %>% filter(df.casy$sample_type == "ABB")
df.casy.bbb = df.casy %>% filter(df.casy$sample_type == "BBB")

df.casy.abb$counts = df.casy.abb$counts*(10010*2*10^3)/(200*10)
df.casy.bbb$counts = df.casy.bbb$counts*(10010*50*10^3)/(200*10)

# Find the mean per channel per BBB or ABB (removing technical error)
# Joining ABB and BBB tables
df.casy.abb.m = data.frame(df.casy.abb %>% 
                            group_by(channel) %>%
                            summarise_at(vars(counts),
                                         list(mean, sd)))
colnames(df.casy.abb.m) = c("channel", "mean", "sd")
df.casy.abb.m$sample_type = "ABB"

df.casy.bbb.m = data.frame(df.casy.bbb %>% 
                            group_by(channel) %>%
                            summarise_at(vars(counts),
                                         list(mean, sd)))
colnames(df.casy.bbb.m) = c("channel", "mean", "sd")
df.casy.bbb.m$sample_type = "BBB"
df.casy.bb.m = rbind(df.casy.abb.m, df.casy.bbb.m)
df.casy.bb.m = left_join(df.design, df.casy.bb.m, by = "channel")


#Plot BBB and ABB Cell Counts per channel
ggplot(df.casy.bb.m, aes(x = as.factor(channel), y = mean/(10^6), fill = sample_type)) +
  geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=(mean-sd)/(10^6), ymax=(mean+sd)/(10^6)), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_manual("legend", values = c("BBB" = "#00BBDB", "ABB" = "#F066EA")) +
  labs(x = "Channel", y = "Million cells / mL",
       title = "Cell counts in BBB and ABB samples") +
    guides(fill=guide_legend(title="Sample type")) +
  theme_bw()

```

### Calculate broken cells per channel

```{r}
df.casy.bbb.c = df.casy.bb.m %>% filter(df.casy.bb.m$sample_type == "BBB")
df.casy.abb.c = df.casy.bb.m %>% filter(df.casy.bb.m$sample_type == "ABB")
df.casy.bbb.c$broken_cells = df.casy.bbb.c$mean - df.casy.abb.c$mean
df.casy.bbb.c$lysis_eff = df.casy.bbb.c$broken_cells/df.casy.bbb.c$mean

# Plot Broken cells in each channel
ggplot(df.casy.bbb.c, aes(x = as.factor(channel), y = broken_cells, fill = purpose)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual("legend", values = c("control" = "#00BBDB", "treatment" = "#F066EA")) +
  labs(x = "Channel", y = "Million broken cells / mL",
       title = "Broken cells in each channel") +
    guides(fill=guide_legend(title="Sample type")) +
  theme_bw()
```

# Lysis Efficiency

```{r}
ggplot(df.casy.bbb.c, aes(x = as.factor(channel), y = lysis_eff, fill = purpose)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual("legend", values = c("control" = "#00BBDB", "treatment" = "#F066EA")) +
  labs(x = "Channel", y = "Lysis efficiency (%)",
       title = "Lysis efficiency per channel") +
    guides(fill=guide_legend(title="Sample type")) +
  scale_y_continuous(labels = scales::percent) +
  theme_bw()
```

### Ploidy Determination.
# Calculate [DNA]/broken_cell for each
```{r}

df.qpcr.smpl.c = data.frame(df.qpcr.smpl %>% 
                            group_by(channel) %>%
                            summarise_at(vars(nDNA_fragments),
                                         list(mean, sd)))
colnames(df.qpcr.smpl.c) = c("channel", "mean", "sd")
df.qpcr.smpl.c$chr_cell = df.qpcr.smpl.c$mean*10^3/df.casy.bbb.c$broken_cells


# Join tables
df.qpcr.smpl.c = left_join(df.qpcr.smpl.c, df.design, by = "channel")
df.chr.cell = data.frame(df.qpcr.smpl.c %>%
                           group_by(purpose) %>%
                           summarise_at(vars(chr_cell),
                                        list(mean, sd)))
colnames(df.chr.cell) = c("purpose", "mean", "sd")
df.chr.cell$Phosphate_mM = c(0.23, 0.0115)
```

# Normality test using Shapiro test for small sample size (min 3, max 5000) https://stats.stackexchange.com/questions/1645/appropriate-normality-tests-for-small-samples
# if the p-value is less than α =.05, there is sufficient evidence to say that the sample does not come from a population that is normally distributed.
```{r}
# Split lists according to conditions, assign the nonconformities to 999
ls_dna_density_treatment <- ifelse(df.qpcr.smpl.c$purpose == "treatment",df.qpcr.smpl.c$chr_cell, 999)

ls_dna_density_control <- ifelse(df.qpcr.smpl.c$purpose == "control",df.qpcr.smpl.c$chr_cell, 999)

# Remove the nonconformities
ls_dna_density_treatment[ ls_dna_density_treatment != 999]
ls_dna_density_control[ ls_dna_density_control != 999]

#Normality test. p<0.05
shapiro.test(ls_dna_density_treatment) #p-value = 0.00416, non-normally distributed
shapiro.test(ls_dna_density_control) #p-value = 0.004231, non-normally distributed
```

# Time to find out if the DNA/cell are different for Treatment and Control
# https://www.statology.org/mann-whitney-u-test-r/
# the p-value is less than α =.05, there is sufficient evidence to reject the null hypothesis, meaning treatment and control are different
# one-tailed test, because we wanna test if less phosphate leads to less ploidy

```{r}
wilcox.test(ls_dna_density_treatment, ls_dna_density_control, alternative = "less")

# p-value = 0.2471. Null hypothesis is not rejected. There's no significant difference between treatment and control. 

```


# Todo: change means to medians +- 95% CI
```{r}
ggplot(data=df.chr.cell, aes(x=as.factor(Phosphate_mM), y= mean, fill = as.factor(Phosphate_mM))) +
  geom_bar(stat="identity", position=position_dodge()) +
  labs(y = expression(paste("chr x cell"^-1)), x = "[Phosphate] (mM)",
       title = "Ploidy level across experimental conditions",
       subtitle = "Bars show mean value per condition and errors indicate 95% CI") +
  scale_fill_manual("legend", values = c("0.23" = "#00BBDB", "0.0115" = "#F066EA")) +
  geom_errorbar(aes(ymin=mean-qnorm(0.95)*sd/sqrt(3), ymax=mean+qnorm(0.95)*sd/sqrt(3)), width=.2,
                 position=position_dodge(.9)) +
  scale_y_continuous(breaks = c(0,5,10)) +
  theme_classic() +
  theme(legend.position="null")


ggplot(data=df.chr.cell, aes(x=df.chr.cell$purpose, y= mean, fill=df.chr.cell$purpose)) +
  geom_boxplot(varwidth = TRUE, notch=TRUE) +
  geom_errorbar(aes(ymin=mean-qnorm(0.95)*sd/sqrt(3), ymax=mean+qnorm(0.95)*sd/sqrt(3)), width=.2,
                 position=position_dodge(.9)) + 
  labs(y = expression(paste("DNA Concentration per cell, [DNA]/cell")), x = "Conditions",
       title = "Ploidy level across experimental conditions",
       subtitle = "Bars show mean value per condition and errors indicate 95% CI") +
  scale_fill_manual("legend", values = c("0.23" = "#00BBDB", "0.0115" = "#F066EA")) +
  scale_y_continuous(breaks = c(0,5,10)) +
  theme_classic() +
  theme(legend.position="null")
```

